---
title: "Predicting the Quality of Weight-Lefting Exercises With Machine Learning in R"
author: "Srawan Kumar Srimant"
date: "May 16, 2017"
output: html_document
---

## Predicting the Quality of Weight-Lefting Exercises With Machine Learning in R

Human Activity Recognition provides the excecise done by different athlete in correct as well as incorrect manner.
The Training dataset provides a variable named Classes of Factor king used to Rate the Quality of the excecise.
Based on the model developed in the Training set the quality of the excecise has to be determined in the Test Set
for 20 different observations.

## Loading the required packages

```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(knitr)
library(randomForest)
```


## Data Import
```{r}
Training <-read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
Validate  <-read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
```

## Partition Training Set
```{r}
inTrain <- createDataPartition(Training$classe, p=0.6, list=FALSE)
Training <- Training[inTrain, ]
Testing <- Training[-inTrain, ]
dim(Training); dim(Testing)
```


## Data Cleaning
## Removing Near Zero Variables
```{r}
nzv <- nearZeroVar(Training, saveMetrics=TRUE)
MyTraining <- Training[,nzv$nzv==FALSE]

nzv <- nearZeroVar(Testing, saveMetrics=TRUE)
MyTesting <- Testing[,nzv$nzv==FALSE]
```

## Removing NA variables from Training Data Set when occurance of NA is more that 60% of the total (19622*60% = 11774)

```{r}
MyTraining1 <-MyTraining[colSums(!is.na(MyTraining)) > 11774]
```

##Remove the first column of the myTraining data set

```{r}
myTraining2 <- MyTraining1[c(-1)]

```

## Removing NA variables from Testing  Data Set when occurance of NA is more that 60% of the total (20*60% = 2825)

```{r}
MyTesting1  <-MyTesting[colSums(!is.na(MyTesting)) > 2825]

MyTesting2 <- MyTesting1[c(-1)]

dim(myTraining2)
dim(MyTesting2)

```


## Model based on Decision Tree Plot.

```{r}
set.seed(12345)
Model_DT <- rpart(classe ~ ., data=myTraining2, method="class")
rpart.plot(Model_DT)

```

## Creating Confusion Matrix on MyTesting2 Dataset

```{r}
Prediction_DT <- predict(Model_DT, MyTesting2, type = "class")
confusionMatrix_tree <- confusionMatrix(Prediction_DT, MyTesting2$classe)
confusionMatrix_tree
plot(confusionMatrix_tree$table, col = confusionMatrix_tree$byClass, main = paste("Decision Tree Confusion Matrix: Accuracy =", round(confusionMatrix_tree$overall['Accuracy'], 4)))

```

## Model based on Random Forest.

```{r}
set.seed(12345)
Model_RF <- randomForest(classe ~ ., data=myTraining2)
Prediction_RF <- predict(Model_RF, MyTesting2, type = "class")
confusionMatrix_RF <- confusionMatrix(Prediction_RF, MyTesting2$classe)
confusionMatrix_RF
plot(Prediction_RF)
plot(confusionMatrix_RF$table, col = confusionMatrix_tree$byClass, main = paste("Random Forest Confusion Matrix: Accuracy =", round(confusionMatrix_RF$overall['Accuracy'], 4)))
```

## Predicting Results on the Validation Data based on Random Forest as its accuracy is better than Decision Tree

```{r}
Req_variables <- colnames(myTraining2[, -58])   # remove the classe column
my_Validate <- Validate[Req_variables]    # allow only variables in validate that are also in myTraining


Predict_validate <- predict(Model_RF, data=my_Validate,type = "class")

```


